---
title: "Pelagic stocks (ICES): TAC simulations"
subtitle: "Read in stocks, updates, and simulates advice"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
# Knit → HTML under report/html/. For PDF/Word see report/knit_to_report.R
# (knit_to_report(..., format = "pdf"|"docx") or render_report(...)).
knit: (function(inputFile, encoding) { e <- new.env(); sys.source(file.path(dirname(normalizePath(inputFile)), "knit_to_report.R"), envir = e, keep.source = FALSE); e$knit_to_report(inputFile, encoding) })
output:
  html_document:
    self_contained: false
    lib_dir: "html/libs/01_pelagics"
    toc: true
    toc_float: true
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
  word_document:
    toc: true
---

```{r}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)

options(warn=-1) 
```

```{r paths}
# Change once if the project moves; package code lives in R/.
projectRoot = "C:/active/Resilience"
source(file.path(projectRoot, "report/ensureBimResilience.R"))
ensureBimResilience(projectRoot)
list2env(resiliencePaths(projectRoot), envir = environment())
```

# Purpose

This file runs the **pelagic** workflow in three stages: 
  **(i)** get chosen stocks, 
  **(ii)** update to the current year, 
  **(iii)** simulate future TACs. 
  
Summaries and discussion are found in `report/00_report.Rmd`.

**Principal outputs:** `data/om/pel-stks.RData`, `data/om/pel-om.RData`, `data/TAC/pel-fct.RData`, `data/TAC/csv/pel-f.csv`.
---

```{r libraries}
loadReportLibraries()
iFig = 0
```

```{r sag}
list2env(
  loadSagBundle(
    projectRoot = projectRoot,
    fishery = "Pelagics",
    sagYears = 2025,
    sdgYear = 2025,
    dropHomEarly = TRUE),
  envir = environment())
```

# (i) Get chosen stocks

Starting FLStocks come from the packaged snapshot (`loadShippedStocks("pelagics")`).
Rebuild from local SS3 / SAM / `icesdata` only when assessments change, then run
`source("data-raw/ship_stocks.R")` to refresh `inst/extdata/`.


```{r pelagics}
# Set TRUE to rebuild from local SS3/SAM (needs data/inputs/), then re-ship.
rebuildPelStks = FALSE

if (isTRUE(rebuildPelStks)) {
  data(icesdata)
  load(file.path(dirIces, "Mac_FLStock_2025.RData"))
  load(file.path(dirSAM, "whbSAM.RData"))
  bocSS = readSS3(file.path(dirSS, "boc.27.6-8"))
  homSS = readSS3(file.path(dirSS, "hom.27.2a3a4a5b6a7a-ce-k8"))
  stks = FLStocks(
    "boc.27.6-8"                    = bocSS[["FLStock"]],
    "hom.27.2a3a4a5b6a7a-ce-k8"     = homSS[["FLStock"]][, -1],
    "mac.27.nea"                    = mac,
    "whb.27.1-91214"                = SAM2FLStock(object = fit))
  for (id in names(stks)) name(stks[[id]]) = id
  rm(fit)
  dir.create(dirOM, recursive = TRUE, showWarnings = FALSE)
  save(stks, file = file.path(dirOM, "pel-stks.RData"))
} else {
  stks = loadShippedStocks("pelagics")
  dir.create(dirOM, recursive = TRUE, showWarnings = FALSE)
  save(stks, file = file.path(dirOM, "pel-stks.RData"))
}
```



```{r, plot}
p1=plot(stks)+
  facet_grid(qname~stock,scale="free")+
  theme(legend.position="bottom")

p2=ggplot(melt(subset(ts,sid%in%names(stks))[,-c(2,8:9)],id=c("sid","year")))+
  geom_line(aes(year,value,col=sid))+
  facet_grid(variable~sid,scale="free")+
  theme(legend.position="bottom")

pelTS=rbind(cbind(Source="SAG",p2$data),
            transmute(p1$data,Source="FLStock",year=year,sid=stock,variable=qname,value=data))
pelTS=merge(pelTS,melt(transmute(rfpts,sid=sid,SB=MSYBtrigger,F=FMSY)),by=c("sid","variable"),all.x=TRUE)

names(pelTS)=gsub("\\.x","",     names(pelTS))
names(pelTS)=gsub("\\.y",".rfpt",names(pelTS))
pelTS[is.na(pelTS$value.rfpt),"value.rfpt"]=1

p3=ggplot(pelTS)+
  geom_line(aes(year,value/value.rfpt,col=Source))+
  facet_grid(variable~sid,scale="free")+
  theme(legend.position="bottom")

p3
```

# (ii) Update to current year

```{r update}
updt = merge(pelTS, ldply(stks, function(x) unlist(dims(x)["maxyear"]), .id = "sid"))
updt = subset(updt, year >= maxyear & Source == "SAG")

ctc = dlply(subset(updt, variable == "C"), .(sid), with, {
  as.FLQuant(data.frame(data = value, year = year))
})
rec = dlply(subset(updt, variable == "Rec"), .(sid), with, {
  window(as.FLQuant(data.frame(data = value, year = year)), end = 2027)
})
rec = llply(rec, function(x) { x[is.na(x)] = median(x[!is.na(x)]); x })

# Stock-specific prep before the shared advice bridge.
id = "boc.27.6-8"
stks[[id]] = FLasher::fwd(
  fwdWindow(stks[[id]], end = 2027),
  sr = rec[[id]],
  catch = ctc[[id]][, ac(2023:2024)])

id = "mac.27.nea"
stks[[id]] = window(stks[[id]], end = 2024)

# Catch years to bridge each stock from assessment maxyear through advice.
bridge = list(
  "boc.27.6-8"                = 2025:2026,
  "hom.27.2a3a4a5b6a7a-ce-k8" = 2025:2026,
  "mac.27.nea"                = 2024:2025,
  "whb.27.1-91214"            = 2025:2026
)
# Trim after bridge so fwdWindow leftovers (NA stock.n) are not carried into
# long-term forecasts. Mac’s last advice catch year is 2025.
endYr = c(
  "boc.27.6-8"                = 2026,
  "hom.27.2a3a4a5b6a7a-ce-k8" = 2026,
  "mac.27.nea"                = 2025,
  "whb.27.1-91214"            = 2026
)

for (id in names(bridge)) {
  x = FLasher::fwd(
    fwdWindow(stks[[id]], end = 2027),
    sr = rec[[id]],
    catch = advice[[id]][, ac(bridge[[id]])]
  )
  x = window(x, end = endYr[[id]])
  stks[[id]] = x
}

stks = FLStocks(stks)
for (id in names(stks)) name(stks[[id]]) = id
```


## Equilibrium and regimes (inputs to TAC scenarios)

`calcEq` per stock, regime detection on stock--recruit residuals, and `recCurrent` for the Regime TAC row. **Result:** `pel-om.RData` and `ggPelRecResidual`.

```{r equilibrium}
list2env(
  calcEqAndRegimes(stks = stks, sid = sids$sid, rfpts = rfpts),
  envir = environment())
save(stks, eqs, srs, rds, ts, rfpts, advice, recCurrent, recRegime,
     file = file.path(dirOM, "pel-om.RData"))
```

```{r recruitment-residuals, fig.height=10, fig.width=14}
ggPelRecResidual = plotRecResiduals(srs, rds)
print(ggPelRecResidual)
```

**Figure `r iFig=iFig+1; iFig`** Pelagic recruitment residuals (fitted stock--recruitment deviations); shaded polygons indicate regime segments.

# (iii) Simulate future TACs

Short-term catch scenarios via `simTAC()` / `saveSimTacResults()` (`fcstCtrl` + FLasher). **Result:** `pel-fct.RData`, `csv/pel-f.csv`, scenario `FLStock` lists.

```{r forecasts}
# Use in-memory objects from equilibrium (do not reload pel-om.RData —
# that file may be stale relative to this knit). For an interactive restart,
# load(file.path(dirOM, "pel-om.RData")) before running this chunk.
assertAfterEquilibrium()

list2env(
  simTAC(
    stks       = stks,
    sid        = sids$sid,
    eqs        = eqs,
    srs        = srs,
    recCurrent = recCurrent,
    rfpts      = rfpts),
  envir = environment())

list2env(
  saveSimTacResults(
    fcstStks  = fcstStks,
    rfpts     = rfpts,
    ctcAdvice = advice,
    dirTAC    = dirTAC,
    prefix    = "pel"),
  envir = environment())

list2env(unpackFcstScenarios(fcstStks), envir = environment())
```

**Figure `r iFig=iFig+1; iFig`** Pelagic forecasts: projected catch under status quo F, Fmsy, and regime recruitment scenarios.

```{r forecast-catch-build}
historical = FLStocks(stks)
ggPelForecast = plotForecastCatch(fct)
print(ggPelForecast)
```

```{r forecast-save-stks, eval=FALSE, include=FALSE}
pelFcstStks = c(fcstStks, list(Historical = historical))
save(pelFcstStks, file = file.path(dirOM, "pel-fcst-stks.RData"))
```

### Optional time-series and projection panels

```{r, eval=FALSE, time-series-strips, fig.height=4, fig.width=14}
tsPlots = buildOptionalTsStrips(stks, eqs, rfpts)
ggPelTsSsb   = tsPlots$ssb
ggPelTsF     = tsPlots$f
ggPelTsCatch = tsPlots$catch
ggarrange(plotlist = ggPelTsSsb,   ncol = 2)
ggarrange(plotlist = ggPelTsF,     ncol = 2)
ggarrange(plotlist = ggPelTsCatch, ncol = 2)
```

```{r, eval=FALSE, stock-plots, fig.height=12}
plotOptionalProjections(fsq, historical, fmsy, rgm, eqs)
```

## Bundled plot objects

```{r, eval=FALSE, save-plot-objects-pelagics, include=FALSE}
saveGroupPlotBundle(
  prefix = "pel",
  recResiduals = ggPelRecResidual,
  forecastCatch = ggPelForecast,
  timeSeries = list(ssb = ggPelTsSsb, f = ggPelTsF, catch = ggPelTsCatch),
  dirOM = dirOM,
  dirPlot = dirPlot)
```

## Summary

Pelagic workflow produces harmonized assessments, regime-aware recruitment diagnostics, multi-scenario short-term catch paths, SAG/advice diagnostic figures, and CSV export **`data/TAC/csv/pel-f.csv`**. Interpretation, methods narrative, and cross-stock conclusions are developed in `report/00_report.Rmd`.

