---
title: "Nephrops functional units:  TAC simulations"
subtitle: "Read in stocks, updates, and simulates advice"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
# Knit → HTML under report/html/. For PDF/Word see report/knit_to_report.R
# (knit_to_report(..., format = "pdf"|"docx") or render_report(...)).
knit: (function(inputFile, encoding) { e <- new.env(); sys.source(file.path(dirname(normalizePath(inputFile)), "knit_to_report.R"), envir = e, keep.source = FALSE); e$knit_to_report(inputFile, encoding) })
output:
  html_document:
    self_contained: false
    lib_dir: "html/libs/03_nephrops"
    toc: true
    toc_float: true
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
  word_document:
    toc: true
---

# Purpose {#purpose}

This report runs the **Nephrops** workflow in three stages: **(i)** get chosen stocks (SAG / functional units), **(ii)** update to the current year (JABBA / biodyn conditioning), **(iii)** simulate future TACs. Methods and synthesis belong in `report/00_report.Rmd`.

**Typical outputs:** `data/om/neph-sag.RData`, `data/om/neph-jabba.RData`, `data/om/neph-bds.RData`, `data/TAC/neph-fct.RData`, `data/TAC/csv/neph-f.csv`.

---

# Introduction

Functional-unit time series are aligned with ICES SAG-style inputs; surplus-production/biodyn-style OMs support JABBA fitting and TAC scenarios. The three stages below match the pelagic/demersal reports; interpretive discussion belongs in `00_report.Rmd`.

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
options(warn = -1)
```

```{r paths}
source(file.path(knitr::current_input(dir = TRUE), "ensureBimResilience.R"))
setupReportPaths()
```

```{r libraries}
loadReportLibraries(extra = c(
  "mpb", "JABBA", "scales", "foreach", "doParallel", "parallel", "forecast"))
iFig = 0
```


# (i) Get chosen stocks

SAG series and reference points for Nephrops functional units in `sid`. **Result:** `nep`, `jabbaInp`, `rfpts`, wide `neph`.

```{r sag}
list2env(
  loadSagBundle(
    projectRoot = projectRoot,
    fishery = "Nephrops",
    sagYears = 2025,
    sdgYear = 2025,
    includeTB = TRUE,
    dropAssYearRows = TRUE,
    requireCatches = FALSE),
  envir = environment())

neph = subset(ts, !is.na(SB))
adviceFile = if (file.exists(adviceCsv)) adviceCsv else shippedAdvicePath(projectRoot)
ctc2026 = read.csv(adviceFile)
save(neph, rfpts, advice, file = file.path(dirOM, "neph-sag.RData"))
```

```{r data-loading}
nephPlotData = subset(merge(neph, rfpts, by = "sid"),
                      !is.na(year) & year >= 2000 & year <= 2025)
nephStockPlotData = subset(nephPlotData, !is.na(SB) & SB >= 0)
nephCatchPlotData = subset(nephPlotData, !is.na(C) | !is.na(landings))
nephFishingPressurePlotData = subset(nephPlotData, !is.na(F) & !is.na(FMSY) & FMSY != 0)
nephFishingPressurePlotData$FRel =
  nephFishingPressurePlotData$F / nephFishingPressurePlotData$FMSY
nephFishingPressurePlotData = subset(
  nephFishingPressurePlotData, !is.na(FRel) & FRel >= 0)

# Shared facet time-series styling used by SAG and JABBA panels below.
nepTs = function(p, ylab, title, free = "free_y", ylim = c(0, NA), hline = NULL) {
  p = p +
    scale_x_continuous(limits = c(2000, 2025), breaks = c(2000, 2010, 2020)) +
    scale_y_continuous(limits = ylim) +
    theme_minimal(12) +
    labs(x = "Year", y = ylab, title = title) +
    facet_wrap(~sid, scale = free, ncol = 3)
  if (!is.null(hline))
    p = p + geom_hline(yintercept = hline, col = "red", linetype = "dashed")
  p
}
```

```{r neph-stock-ts, fig.height=8, fig.width=12}
ggNepStockTs = nepTs(
  ggplot(nephStockPlotData) +
    geom_path(aes(year, SB)) +
    geom_hline(
      aes(yintercept = MSYBtrigger),
      data = subset(nephStockPlotData, !is.na(MSYBtrigger)),
      col = "red", linetype = "dashed"),
  ylab = "Stock",
  title = expression("Stock relative to MSY" ~ italic(B[trigger])),
  free = "free")
ggNepStockTs
```

**Figure `r iFig=iFig+1; iFig`**: Stock biomass time series for Nephrops functional units from ICES assessment data.

```{r neph-catch-ts, fig.height=8, fig.width=12}
ggNepCatch = nepTs(
  ggplot(nephCatchPlotData) +
    geom_path(aes(year, C),
              data = subset(nephCatchPlotData, !is.na(C) & C >= 0),
              col = "blue") +
    geom_path(aes(year, landings),
              data = subset(nephCatchPlotData, !is.na(landings) & landings >= 0),
              col = "darkgreen"),
  ylab = "Landings",
  title = "Catch and Landings")
ggNepCatch
```

**Figure `r iFig=iFig+1; iFig`**: Catch and landings time series for Nephrops functional units.

```{r neph-fishing-pressure-ts, fig.height=8, fig.width=12}
ggNepFishingPressureTs = nepTs(
  ggplot(nephFishingPressurePlotData) +
    geom_path(aes(year, FRel), col = "blue"),
  ylab = "F",
  title = "Fishing Pressure",
  hline = 1)
ggNepFishingPressureTs
```

**Figure `r iFig=iFig+1; iFig`**: Fishing pressure indices for Nephrops functional units.

```{r neph-mean-size-ts, fig.height=8, fig.width=12,eval=FALSE}
#' Plot Mean Size Over Time
#' 
#' @param data Stock data
#' @param ncol Number of columns for faceting (default: 3)
#' @return ggplot object
#' @export
plotMeanSize = function(data, ncol = 3) {
  # Determine which column to use for faceting (sid or .id)
  facet_col = if ("sid" %in% names(data)) "sid" else ".id"
  
  p = ggplot(data) +
    geom_point(aes(year, C/(SB*F)), col = "blue") +
    theme_minimal(12) +
    labs(x = "Year",
         y = "F",
         title = "Mean Size")
  
  # Add faceting with appropriate column
  if (facet_col == "sid") {
    p = p + facet_wrap(~sid, scale = "free_y", ncol = ncol)
  } else {
    p = p + facet_wrap(~.id, scale = "free_y", ncol = ncol)
  }
  
  return(p)
}


# Plot mean size
ggNepMeanSizeTs = plotMeanSize(nep, ncol = 3)+
    scale_y_continuous(limits=c(0,NA))
print(ggNepMeanSizeTs)
```

**Figure `r iFig=iFig+1; iFig`**: Mean size time series for Nephrops functional units. Each panel shows the average size of individuals over time for individual functional units. 

# (ii) Update to current year

JABBA fits and biodyn OMs conditioned through the reporting year (advice bridge where needed).

```{r neph-jabba}
# r:     intrinsic growth rate
# shape: production function shape (fixed)
# psi:   initial depletion (B1/K)
# k:     carrying capacity
# sigma: process error 
# tau:   observation error

fillYr <- function(df, yrCol = "year", xCol = "catch") {
  stopifnot(all(c(yrCol, xCol) %in% names(df)))
  
  df=df[order(df[[yrCol]]), ]
  fullYrs=data.frame(year = seq(min(df[[yrCol]], na.rm = TRUE),
                                   max(df[[yrCol]], na.rm = TRUE),
                                   by = 1))
  
  df2=merge(fullYrs, df, by.x = "year", by.y = yrCol, all.x = TRUE)
  df2=df2[order(df2$year), ]
  
  x=df2[[xCol]]
  
  for (i in which(is.na(x))) {
    left=if (i > 1) max(which(!is.na(x[1:(i - 1)])), na.rm = TRUE) else NA_integer_
    right=if (i < length(x)) {
      w=which(!is.na(x[(i + 1):length(x)]))
      if (length(w)) min(w) + i else NA_integer_
    } else NA_integer_
    
    if (is.finite(left) && is.finite(right)) {
      x[i]=mean(c(x[left], x[right]))
    }
  }
  
  df2[[xCol]]=x
  df2}

inp=transmute(neph,sid  =sid,
                   year =year,
                   catch=ifelse(is.na(C),landings+discards, C),
                   index=SB)
inp=merge(inp,rfpts)

ncore=detectCores() - 4  
cl=makeCluster(ncore)
registerDoParallel(cl)

jbNep=foreach(id=sids$sid, .combine=list, 
                .maxcombine=length(rfpts$sid), 
                .packages  =c("dplyr","JABBA","FLBRP","mpb","reshape","FLRebuild","plyr","zoo")) %dopar% {

          dat  =fillYr(subset(inp,sid==id&year<2025))

          pr   =FLPar(c(r=0.4,shape=0.4,psi=0.8,current=tail(dat$index,1)/dat$MSYBtrigger[1]*0.6))
          
          if(is.na(pr["current"])) pr["current"]=0.6

          pr.sd=pr/pr*0.3


          catch=dat[,c("year","catch")]

          Index=dat[,c("year","index")]
          #auxiliary=dat[,c("year","ffmsy")]

          
          rtn=tryIt(FLRebuild::jabba(catch, pr, pr.sd=pr.sd, assessment=id, 
                                      fixed.obsE=0.3,index=Index, current="b"))
          
          rtn[["biodyn"]] = tryIt(mpb:::jabba2biodyn(rtn[[2]]))
          
          rtn}

stopCluster(cl)

id=9
plot(jbNep[[id]][[3]])+
   geom_flpar(data=FLPars(Stock  =FLPar("Bmsy"=refpts(jbNep[[id]][[3]])["bmsy",drop=T]),
                          Harvest=FLPar("Fmsy"=refpts(jbNep[[id]][[3]])["fmsy",drop=T]),
                          Yield  =FLPar("MSY" =refpts(jbNep[[id]][[3]])[ "msy",drop=T])),
              x=dims(jbNep[[id]][[3]])$minyear+1)

names(jbNep)=sids$sid

save(jbNep,file=file.path(dirOM, "neph-jabba.RData"))
```

```{r neph-jabba-stock-ts, fig.height=8, fig.width=12}
jbN = ldply(jbNep, function(x) {
  if (is.null(x)) return(NULL)
  tryIt(model.frame(
    FLQuants(
      x[[3]],
      Stock   = function(x) stock(x)   %/% refpts(x)["bmsy"],
      Harvest = function(x) harvest(x) %/% refpts(x)["fmsy"],
      Yield   = function(x) catch(x)   %/% refpts(x)["msy"]
    ),
    drop = TRUE))
}, .id = "sid")

ggNepJabbaStockTs = nepTs(
  ggplot(jbN) + geom_path(aes(year, Stock)),
  ylab = "Stock",
  title = expression("Stock relative to MSY" ~ italic(B[trigger])),
  free = "free",
  hline = 1)
ggNepJabbaStockTs
```

**Figure `r iFig=iFig+1; iFig`**: JABBA stock biomass relative to BMSY by functional unit.

```{r neph-jabba-yield-ts, fig.height=8, fig.width=12}
ggNepJabbaYieldTs = nepTs(
  ggplot(jbN) + geom_path(aes(year, Yield), col = "blue"),
  ylab = "Yield / MSY",
  title = "Catch relative to MSY",
  hline = 1)
ggNepJabbaYieldTs
```

**Figure `r iFig=iFig+1; iFig`**: JABBA yield relative to MSY by functional unit.

```{r neph-jabba-harvest-ts, fig.height=8, fig.width=12}
ggNepJabbaHarvestTs = nepTs(
  ggplot(jbN) + geom_path(aes(year, Harvest), col = "blue"),
  ylab = "F",
  title = "Fishing Pressure",
  hline = 1)
ggNepJabbaHarvestTs
```

**Figure `r iFig=iFig+1; iFig`**: JABBA fishing pressure relative to FMSY by functional unit.

```{r neph-process-error-ts}
toProcessErrorSeries <- function(y) {
  data.frame(
    year = as.numeric(dimnames(y[["timeseries"]])[[1]]),
    pe = as.numeric(y[["timeseries"]][,"mu","procB"]))}

jbPE=ldply(jbNep, function(x) toProcessErrorSeries(x[["fit"]]),.id="sid")
rds=ddply(jbPE, .(sid), with, rod(as.FLQuant(data.frame(year=year,data=pe))), .id = "sid")

recRegime=ddply(rds, .(sid, regime), with, 
                data.frame(data=exp(mean(data))), .id="sid")

# Get current (most recent) regime
recCurrent=ddply(recRegime, .(sid), with, 
                 data.frame(regime=tail(data, 1)))

save(jbN, jbPE, rds, recRegime, recCurrent, file = file.path(dirOM, "neph-jabba-outputs.RData"))

ggNepProcResiduals=ggplot(jbPE)+
  facet_wrap(~sid,scale="free",ncol=2)+
  geom_polygon(aes(year,data,group=regime),
               fill="lavender",col="blue",
               lwd=.25,alpha=.5,
               data=rds)+
  geom_line( aes(year,pe),col="grey75")+
  geom_point(aes(year,pe),col="grey75",fill="grey25",shape=21)+
  xlab("Year")+ylab("Residual")+
  theme_bw(16)+
  theme(axis.text.x=element_text(angle=-30),legend.position="bottom")

ggNepProcErrorTs = nepTs(
  ggplot(jbPE) + geom_path(aes(year, pe), col = "blue"),
  ylab = "",
  title = "Process Error",
  ylim = c(NA, NA),
  hline = 0)
ggNepProcResiduals
ggNepProcErrorTs
```

```{r neph-om}
## last catch year, if < 2026 then provide advice

bds=llply(jbNep,function(x) x[["biodyn"]])
flag=!laply(bds,is.null)
bds=mpb:::biodyns(bds[flag])


ctcAdvice=melt(read.csv(adviceFile, stringsAsFactors=FALSE))
ctcAdvice=ctcAdvice[ctcAdvice$sid%in%names(jbNep),]
ctcAdvice=transmute(ctcAdvice,sid=sid,data=value,year=substr(variable,2,5))
ctcAdvice=FLQuants(dlply(ctcAdvice,.(sid), with, as.FLQuant(data.frame(year=year,data=data))))
ctcAdvice=ctcAdvice[names(bds)]

bds=buildNephOps(bds,
                 advice           = ctcAdvice,
                 sqYears          = 2022:2024, #3
                 projectionYears  = 2027:2040,
                 processErrorSd   = 0.15,
                 processErrorB    = 0.3)

save(bds, file = file.path(dirOM, "neph-bds.RData"))

# plot(mcf(FLQuants(bd,Catch=function(x) catch(x),
#                       Stock=function(x) stock(x),
#                       "Harvest~Rate"=function(x) harvest(x))), iter=5)+
#     scale_y_continuous(limits=c(0,NA))+
#   xlab("Year") + ggtitle("Single Stock: nep.fu.11")
```

# (iii) Simulate future TACs

Scenario catch trajectories and standard TAC CSV (`neph-f.csv`).

```{r neph-fct}
fct=ldply(bds, function(x) ldply(x, function(y) 
  model.frame(FLQuants(y, Catch=FLCore::catch,Stock=stock,F=harvest),drop=TRUE),.id="Scenario"),.id="sid")

save(fct, file = file.path(dirTAC, "neph-fct.RData"))

# Harmonised forecast CSV (same schema as pel-f.csv / dem-f.csv / iccat-f.csv).
tacCsvDir = file.path(dirTAC, "csv")
if (!dir.exists(tacCsvDir)) dir.create(tacCsvDir, recursive = TRUE, showWarnings = FALSE)
tacNephF = buildForecastTacCsv(
  fct = fct,
  triggerBySid = subset(rfpts, sid %in% unique(as.character(fct$sid)))[, c("sid", "MSYBtrigger")],
  triggerCol = "MSYBtrigger",
  minYear = 2021
)
write.csv(tacNephF, file = file.path(tacCsvDir, "neph-f.csv"), row.names = FALSE)
```


```{r plot-nephrops, fig.width=12, fig.height=8}
ids=names(bds)
if (length(ids) == 0) stop("No Nephrops operating models available in bds.")

ctc=FLQuants(lapply(ids, function(id) {
  bd=bds[[id]][[1]]
  catch(bd) / refpts(bd)["msy"]
}))
names(ctc)=ids

ggNepForecastCatch=plot(ctc)+
    geom_hline(aes(yintercept=1),col="grey")+
    facet_wrap(~qname,scale="free_y",ncol=4)+
    scale_y_continuous(limits=c(0,NA))+
    #geom_text(aes(x=1950, y=2, label=names(cat1[qname])), 
    #        hjust=0, vjust=1, size=3, color="black")+
    labs(title="Cat 1: Catch & SSB",   
         x    ="Year",
         y    =expression(Catch/MY)) +
  theme_minimal(base_size = 12) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    #strip.text = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major = element_line(color = "grey90"),
    plot.title = element_text(face = "bold", size = 14),
    plot.subtitle = element_text(size = 12, color = "grey30"),
    panel.spacing = unit(0, "lines"))

ggNepForecastCatch
```

**Figure `r iFig=iFig+1; iFig`**: Catch relative to MSY for all Nephrops functional units. Values above 1 indicate catch exceeding MSY reference levels. Each panel shows a different functional unit.


```{r devs}
df=subset(neph ,sid=="nep.fu.22")

# Fit ARIMA(1,1,0) with drift
fit=auto.arima(df$landings)

# Simulate 100 series of length nyears
set.seed(123)
sims=replicate(100, simulate(fit, nsim = nrow(df)))

sims=as.FLQuant(c(sims),dimnames=list(year=2023:2047,iter=seq(100)))
```


```{r tac-nephrops}
ids=names(bds)
if (length(ids) == 0) stop("No Nephrops operating models available in bds.")

TACNephrops=ldply(bds, function(x) transmute(as.data.frame(catch(x[[1]])),
                                        iter = iter, year = year, TAC = data))
write.csv(
  subset(TACNephrops, year >= 2021 & year < 2041),
  file = file.path(dirTAC, "csv", "neph-tac-sq.csv"),
  row.names = FALSE)
```

```{r save-nephrops-plot-objects, include=FALSE}
dir.create(dirOM, recursive = TRUE, showWarnings = FALSE)
dir.create(dirPlot, recursive = TRUE, showWarnings = FALSE)

nephPlots = list(
  check = list(
    catch = ggNepCatch,
    ssb = ggNepStockTs,
    f = ggNepFishingPressureTs
  ),
  cfCatch = ggNepCatch,
  recResiduals = ggNepProcResiduals,
  forecastCatch = ggNepForecastCatch
)
save(nephPlots, file = file.path(dirOM, "neph_plot.RData"))

ggPlots03Nephrops = list(
  stockTs = ggNepStockTs,
  catchLandingsTs = ggNepCatch,
  fishingPressureTs = ggNepFishingPressureTs,
  jabbaStockTs = ggNepJabbaStockTs,
  jabbaYieldTs = ggNepJabbaYieldTs,
  jabbaHarvestTs = ggNepJabbaHarvestTs,
  procResiduals = ggNepProcResiduals,
  procErrorTs = ggNepProcErrorTs
)

save(ggPlots03Nephrops, file = file.path(dirPlot, "neph_plot_html.RData"))
```

# Summary of results {#summary-of-results}

This workflow yields per-functional-unit diagnostics, JABBA/biodyn panels, and TAC paths. Key artefacts: **`data/om/neph-jabba.RData`**, **`neph-jabba-outputs.RData`**, **`neph-bds.RData`**, **`neph_plot.RData`**, **`data/TAC/csv/neph-f.csv`**, **`neph-tac-sq.csv`**. Synthesis belongs in `report/00_report.Rmd`.

Cross-unit interpretation, advice policy discussion, and climate or mixed-fishery context should be developed in the synthesis report rather than duplicated here.

## References


