---
title: "North Atlantic albacore (ICCAT): analysis and results"
subtitle: "SS3 stock-only workflow (parallel to ICES pelagic pipeline in `01_pelagics.Rmd`)"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
# Knit → HTML under report/html/. For PDF/Word see report/knit_to_report.R
# (knit_to_report(..., format = "pdf"|"docx") or render_report(...)).
knit: (function(inputFile, encoding) { e <- new.env(); sys.source(file.path(dirname(normalizePath(inputFile)), "knit_to_report.R"), envir = e, keep.source = FALSE); e$knit_to_report(inputFile, encoding) })
output:
  html_document:
    self_contained: false
    lib_dir: "html/libs/04_iccat"
    toc: true
    toc_float: true
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
  word_document:
    toc: true
---

# Purpose

This report runs the **ICCAT albacore** workflow in three stages: **(i)** get the chosen stock (`alb-n`), **(ii)** update with advice catches, **(iii)** simulate future TACs. Methods and multi-stock discussion belong in `report/00_report.Rmd`.

**Outputs:** **`alb_plot.RData`**, **`iccat-fct.RData`**, **`csv/iccat-f.csv`**.

---


```{r}
knitr::opts_chunk$set(warning = FALSE, 
                      message = FALSE)

options(warn=-1) 
```

```{r paths}
# Change once if the project moves; package code lives in R/.
projectRoot = "C:/active/Resilience"
source(file.path(projectRoot, "report/ensureBimResilience.R"))
ensureBimResilience(projectRoot)
list2env(resiliencePaths(projectRoot), envir = environment())
```


```{r libraries}
loadReportLibraries(extra = "ggpubr")
iFig = 0
```

# (i) Get chosen stocks

Starting albacore OM comes from the packaged snapshot (`loadShippedStocks("iccat")`).
Rebuild from local SS3 only when the ICCAT assessment changes, then run
`source("data-raw/ship_stocks.R")` to refresh `inst/extdata/`.

```{r data}
sids = subset(bimSids(), fishery == "ICCAT")
sid = sids$sid

# Set TRUE to rebuild from local SS3 (needs data/inputs/SS/alb-n/), then re-ship.
rebuildAlbOm = FALSE

if (isTRUE(rebuildAlbOm)) {
  alb = readSS3(file.path(dirSS, "alb-n"))
} else {
  alb = loadShippedStocks("iccat")
}

stks = FLStocks("alb-n" = alb$FLStock)
sr = alb$SRR
rfs = alb$refpts

maxYr = ldply(stks, function(x) data.frame(maxyear = dims(x)[["maxyear"]]), .id = "sid")
ts = ldply(stks, function(x)
  model.frame(FLQuants(x, Catch    = FLCore::catch,
                          SSB      = FLCore::ssb,
                          F        = FLCore::fbar,
                          Recruits = FLCore::rec), drop = TRUE))
names(ts)[1] = "sid"
ts = merge(ts, maxYr)
```

# (ii) Update to current year

Advice catches from `advice.csv` bridge the assessment into the projection window.

```{r, advice}
adviceFile = if (file.exists(adviceCsv)) adviceCsv else shippedAdvicePath(projectRoot)
advice = melt(read.csv(adviceFile, stringsAsFactors = FALSE))
advice = advice[advice$sid %in% sid, ]
advice = transmute(advice, sid = sid, data = value, year = substr(variable, 2, 5))
advice = FLQuants(dlply(advice, .(sid), with, as.FLQuant(data.frame(year = year, data = data))))
```

## Stock--recruit fit and recruitment residuals

**Done:** Fit Beverton--Holt SRR on the accepted assessment series; plot residuals for regime-style interpretation in synthesis. **Result:** `ggAlbRecResidual`.

```{r equilibrium-and-residuals, fig.height=6, fig.width=10}
eq=FLBRP(stks[[sid]])
sr=fmle(as.FLSR(stks[[sid]], model="bevholt"), control=list(silent=TRUE))

model(eq) =bevholt()$model
params(eq)=params(sr)

srdf=model.frame(FLQuants(sr,hat=predict,
                            rec=FLCore::rec,
                            rsd=residuals),drop=TRUE)

ggAlbRecResidual=ggplot(cbind(sid="alb-n",srdf))+
  facet_wrap(~sid, scale="free", ncol=1)+
  geom_line(aes(year, rsd), col="grey75")+
  geom_point(aes(year, rsd), col="grey75", fill="grey25", shape=21)+
  xlab("Year")+ylab("Residual")+
  theme_bw(12)
print(ggAlbRecResidual)
```

## Assessment time-series panels

Catch, SSB, F, and recruitment with SS3 MSY-class reference lines. **Result:** four-panel `ggarrange` and separate ggplot objects saved below.

```{r check, fig.height=8, fig.width=12, results='hide'}
theme_set(theme_minimal()+
            theme(axis.title.y=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks.y=element_blank(),
                  axis.line.y=element_blank()))

ggAlbCatch=ggplot(ts)+
  facet_grid(sid~., scale="free_y")+
  geom_line(aes(year, Catch))+
  geom_hline(yintercept=as.numeric(rfs["MSY", drop=TRUE]),inherit.aes=FALSE)+
  labs(title="Catch")+xlab("Year")

ggAlbSsb=ggplot(ts)+
  facet_grid(sid~., scale="free_y")+
  geom_line(aes(year, SSB))+
  geom_hline(yintercept=as.numeric(rfs["SBMSY", drop=TRUE]),inherit.aes=FALSE)+
  labs(title="SSB")+xlab("Year")

ggAlbF=ggplot(ts)+
  facet_grid(sid~., scale="free_y")+
  geom_line(aes(year, F))+
  geom_hline(yintercept=as.numeric(rfs["FMSY", drop=TRUE]),inherit.aes=FALSE)+
  labs(title="F")+xlab("Year")

ggAlbRecruits=ggplot(ts)+
  facet_grid(sid~., scale="free_y")+
  geom_line(aes(year, Recruits))+
  geom_hline(yintercept=as.numeric(rfs["R0", drop=TRUE])*2,inherit.aes=FALSE)+
  labs(title="Recruits")+xlab("Year")

ggarrange(ggAlbCatch, ggAlbSsb, ggAlbF, ggAlbRecruits)
```


# (iii) Simulate future TACs

Forward projections under recruitment multipliers; writes `iccat-fct.RData` and `csv/iccat-f.csv`.

**Done:** Hold recent F forward; compare deterministic paths for **FStatus Quo**, **FStatus Quo 75%**, and **FStatus Quo 50%** (recruitment scaling). **Result:** `fct` + `refPars` in `iccat-fct.RData`, `ggAlbFwdRecMult`.

```{r forecasts, fig.height=8, fig.width=12}
theme_set(theme_minimal()+
            theme(axis.title.y=element_blank(),
                  axis.text.y=element_blank(),
                  axis.ticks.y=element_blank(),
                  axis.line.y=element_blank()))

ftar  =mean(fbar(stks[[sid]])[,ac(dims(stks[[sid]])$maxyear-3:1)])
yrs   =ac(2027:2050)
recDev=FLQuant(1.0,dimnames=list(year=yrs))
ctrl  =fwdControl(year=as.numeric(yrs),
                  quant="f",
                  value=rep(ftar, length(yrs)))

stk=fwdWindow(stks[[sid]], end=max(an(yrs)))

stk=FLasher::fwd(stk,catch=advice[["alb-n"]][,ac(2022:2026)],sr=sr,deviances=recDev)

alb=FLStocks("FStatus Quo"       =FLasher::fwd(stk,control=ctrl,sr=sr,deviances=recDev),
             "FStatus Quo 75%"   =FLasher::fwd(stk,control=ctrl,sr=sr,deviances=recDev*0.75),
             "FStatus Quo 50%"   =FLasher::fwd(stk,control=ctrl,sr=sr,deviances=recDev*0.50))

fct=cbind(sid="alb-n",ldply(alb,   function(x)
    model.frame(FLQuants(x, Catch=FLCore::catch,SSB=ssb,F=fbar,Recruits=rec),drop=TRUE),.id="Scenario"))

refPars=FLPars(
  SB=FLPar("SBMSY"=rfs["SBMSY", drop=TRUE]),
  F=FLPar("FMSY"=rfs["FMSY", drop=TRUE]),
  Yield=FLPar("MSY"=rfs["MSY", drop=TRUE])
)

save(fct, refPars, file = file.path(dirTAC, "iccat-fct.RData"))

tacCsvDir = file.path(dirTAC, "csv")
if (!dir.exists(tacCsvDir)) dir.create(tacCsvDir, recursive = TRUE, showWarnings = FALSE)
sb_msy = suppressWarnings(as.numeric(rfs["SBMSY", drop = TRUE]))
sb_msy = sb_msy[is.finite(sb_msy)]
if (!length(sb_msy)) {
  stop("04_iccat: need finite SBMSY from SS3 reference points (rfs) to compute Btrig for csv/iccat-f.csv")
}
sb_msy = sb_msy[1]
albTrig = data.frame(sid = "alb-n", MSYBtrigger = sb_msy, stringsAsFactors = FALSE)
tacIccatF = buildForecastTacCsv(
  fct = fct,
  triggerBySid = albTrig,
  triggerCol = "MSYBtrigger",
  minYear = 2021
)
write.csv(tacIccatF, file = file.path(tacCsvDir, "iccat-f.csv"), row.names = FALSE)

ggAlbFwdRecMult=plot(alb)+
  geom_flpar(data=refPars, x=min(as.numeric(yrs)))

print(ggAlbFwdRecMult)
```

```{r save-alb-plot-objects, include=FALSE}
plotObjPath=file.path(dirOM, "alb_plot.RData")
if (!dir.exists(dirname(plotObjPath))) dir.create(dirname(plotObjPath), recursive=TRUE, showWarnings=FALSE)
albPlots=list(
  check=list(catch=ggAlbCatch, ssb=ggAlbSsb, f=ggAlbF, recruits=ggAlbRecruits),
  cfCatch=ggAlbCatch,
  recResiduals=ggAlbRecResidual,
  forecastCatch=ggAlbFwdRecMult)

save(albPlots, file=plotObjPath)

dir.create(file.path(dirReport, "figs"), recursive = TRUE, showWarnings = FALSE)
ggPlots04Iccat = list(
  recResiduals=ggAlbRecResidual,
  checkCatch=ggAlbCatch,
  checkSsb=ggAlbSsb,
  checkF=ggAlbF,
  checkRecruits=ggAlbRecruits,
  forecastCatch=ggAlbFwdRecMult
)
save(ggPlots04Iccat, file=file.path(dirReport, "figs", "alb_plot_html.RData"))
```

## Summary

Albacore workflow produces SS3-conditioned diagnostics, recruitment residual plot, recruitment-scaled forward projections, **`data/TAC/csv/iccat-f.csv`** (deterministic scenarios), and packaged ggplot lists for `00_report.Rmd`.
