---
title: "Beamer figures: biological simulations"
subtitle: "Exports PNG panels for `report/beamer/biological_simulations_slides.tex`"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
# Knit → HTML under report/html/. For PDF/Word see report/knit_to_report.R
# (knit_to_report(..., format = "pdf"|"docx") or render_report(...)).
knit: (function(inputFile, encoding) { e <- new.env(); sys.source(file.path(dirname(normalizePath(inputFile)), "knit_to_report.R"), envir = e, keep.source = FALSE); e$knit_to_report(inputFile, encoding) })
output:
  html_document:
    self_contained: false
    lib_dir: "html/libs/05_figs"
    toc: true
    toc_float: true
    code_folding: show
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
  word_document:
    toc: true
---

# Purpose

Build the figures used in the biological-simulations Beamer deck
(`report/beamer/`). PNGs are written next to the `.tex` file.

- **fig1--fig3**: SAG retrospectives (blueMarine)
- **fig4--fig6**: pelagic / demersal / Nephrops forecasts (`data/plot-objects/`)
- **fig_mac_***: mackerel what-ifs from `pel-om.RData` (recruitment levels,
  M shock, random recruitment)

---

```{r setup}
knitr::opts_chunk$set(warning = TRUE, message = TRUE, error = FALSE)
# fail-fast: do not swallow warnings (was options(warn = -1))
```

```{r paths}
source(file.path(dirname(knitr::current_input(dir = TRUE)), "ensureBimResilience.R"))
setupReportPaths()
loadReportLibraries(extra = "ggplot2")

dirBeamer = file.path(dirReport, "beamer")
dir.create(dirBeamer, recursive = TRUE, showWarnings = FALSE)

# SAG snapshot lives in blueMarine (override with BLUEMARINE_DATA if needed).
dirBlueMarine = Sys.getenv(
  "BLUEMARINE_DATA",
  unset = "C:/active/blueMarine/data")
```

```{r libs}
library(ggplot2)
```

```{r themes}
order_stock_names = function(d, stocks, col = "name") {
  d[[col]] = factor(d[[col]], levels = stock_name_levels(stocks))
  d
}

stock_name_levels = function(stocks) as.character(stocks$name)

# Greyscale ramp over assessment years (light → dark).
retro_pal = function(years, cols = c("grey75", "black")) {
  yrs = sort(unique(years))
  setNames(colorRampPalette(cols)(length(yrs)), yrs)
}

theme_beamer_ck = function(base_size = 16, base_family = "sans") {
  theme_bw(base_size = base_size, base_family = base_family) +
    theme(
      plot.background   = element_rect(fill = "transparent", colour = NA),
      panel.background  = element_rect(fill = "white", colour = NA),
      legend.background = element_rect(fill = "transparent", colour = NA),
      legend.key        = element_rect(fill = "transparent", colour = NA),
      panel.grid.minor  = element_blank(),
      panel.grid.major  = element_line(colour = "grey88", linewidth = 0.3),
      panel.border      = element_rect(colour = "grey50", fill = NA, linewidth = 0.4),
      strip.background  = element_rect(fill = "grey85", colour = NA),
      strip.text        = element_text(
        face = "bold", size = rel(1.0), margin = margin(3, 3, 3, 3)),
      axis.title        = element_text(face = "bold", size = rel(1.05)),
      axis.text         = element_text(size = rel(0.9), colour = "grey20"),
      axis.ticks        = element_line(colour = "grey50", linewidth = 0.3),
      axis.text.y       = element_blank(),
      plot.title        = element_text(face = "bold", size = rel(1.15), hjust = 0),
      plot.subtitle     = element_text(size = rel(0.95), colour = "grey30"),
      legend.position   = "none",
      legend.title      = element_text(face = "bold", size = rel(0.95)),
      legend.text       = element_text(size = rel(0.9)),
      plot.margin       = margin(4, 8, 4, 4))
}

theme_set(theme_beamer_ck())

# Advice panels: stock name as a top strip (not a bulky left column),
# with readable y-scales. Does not inherit axis.text.y = blank from theme_beamer_ck.
theme_advice_panel = function(base_size = 12) {
  theme_bw(base_size = base_size) +
    theme(
      plot.background   = element_rect(fill = "white", colour = NA),
      panel.background  = element_rect(fill = "white", colour = NA),
      legend.position   = "none",
      panel.grid.minor  = element_blank(),
      panel.grid.major  = element_line(colour = "grey92", linewidth = 0.25),
      panel.border      = element_rect(colour = "grey55", fill = NA, linewidth = 0.35),
      panel.spacing.y   = unit(8, "pt"),
      strip.background  = element_rect(fill = "white", colour = NA),
      strip.text        = element_text(
        face = "bold", size = rel(1.1), hjust = 0, colour = "grey10",
        margin = margin(t = 2, b = 2)),
      strip.placement   = "outside",
      axis.title        = element_text(face = "bold", size = rel(0.95)),
      axis.title.y      = element_text(margin = margin(r = 4)),
      axis.text         = element_text(size = rel(0.78), colour = "grey25"),
      axis.ticks        = element_line(colour = "grey55", linewidth = 0.25),
      plot.margin       = margin(4, 8, 4, 6))
}

# Compact y labels for absolute biomass; plain decimals for ratios.
.label_ssb_abs = function(x) {
  vapply(x, function(z) {
    if (is.na(z)) return(NA_character_)
    if (abs(z) >= 1e6) sprintf("%.1fM", z / 1e6)
    else if (abs(z) >= 1e3) sprintf("%.0fk", z / 1e3)
    else sprintf("%.2g", z)
  }, character(1))
}

# Retrospective SSB (or SSB/Btrigger) by stock × assessment year.
plot_advice_ssb = function(d, pal, y = "ssb", ylab = "SSB",
                           xlim = c(1980, 2025), hline = NULL,
                           anchor = NULL, status_bands = FALSE,
                           band_at = 1) {
  p = ggplot(d, aes(year, .data[[y]], group = assYr, colour = factor(assYr)))
  if (!is.null(anchor)) {
    p = p + geom_blank(data = anchor, aes(year, ssb), inherit.aes = FALSE)
  }
  # Colourblind-safe bands (Okabe–Ito): blue = carry on (>1), orange = reduce (<1).
  if (isTRUE(status_bands)) {
    p = p +
      geom_rect(
        xmin = -Inf, xmax = Inf, ymin = band_at, ymax = Inf,
        fill = "#56B4E9", alpha = 0.07, inherit.aes = FALSE) +
      geom_rect(
        xmin = -Inf, xmax = Inf, ymin = -Inf, ymax = band_at,
        fill = "#E69F00", alpha = 0.07, inherit.aes = FALSE)
  }
  if (!is.null(hline)) {
    p = p + geom_hline(
      yintercept = hline, colour = "grey20", linetype = 2, linewidth = 0.7)
  }

  y_scale = if (identical(y, "ssb_rel")) {
    scale_y_continuous()
  } else {
    scale_y_continuous(labels = .label_ssb_abs)
  }

  p +
    geom_line(linewidth = 0.55, alpha = 0.9) +
    facet_wrap(~name, ncol = 1, scales = "free_y", strip.position = "top") +
    y_scale +
    coord_cartesian(xlim = xlim) +
    scale_colour_manual(values = pal) +
    labs(x = NULL, y = ylab) +
    theme_advice_panel()
}

save_beamer_fig = function(plot, name, width, height) {
  ggsave(
    filename = file.path(dirBeamer, name),
    plot     = plot,
    width    = width,
    height   = height)
}
```

```{r data}
load(file.path(dirBlueMarine, "interim", "sag_stocks.RData"))
load(file.path(dirBlueMarine, "results", "stock_summary.RData"))
```

```{r sag-ssb, fig.height=10, fig.width=10}
ts_plot = merge(
  sag$ts,
  stock_summary$stocks[, c("sid", "name")],
  by.x = ".id", by.y = "sid")
ts_plot = merge(ts_plot, sag$rfpts, by = c(".id", "assYr"))
ts_plot = order_stock_names(
  subset(ts_plot, is.finite(ssb)),
  stock_summary$stocks)

# Single stock order for fig1--fig3 (do not let subset/anchor drop levels).
stock_order = levels(ts_plot$name)

pal = retro_pal(ts_plot$assYr)

d1 = subset(ts_plot, assYr %in% 2016:2021)
d2 = subset(ts_plot, year >= 1980 & assYr %in% 2016:2026)
d1$name = factor(as.character(d1$name), levels = stock_order)
d2$name = factor(as.character(d2$name), levels = stock_order)

# Shared per-stock y-range so fig1 and fig2 are comparable.
combo = rbind(d1, d2)
rng = do.call(
  rbind,
  lapply(split(combo$ssb, factor(combo$name, levels = stock_order)),
         range, na.rm = TRUE))
rng = data.frame(
  name = factor(rownames(rng), levels = stock_order),
  ymin = rng[, 1],
  ymax = rng[, 2],
  row.names = NULL)

anchor = data.frame(
  year = 1980,
  ssb  = c(rng$ymin, rng$ymax),
  name = factor(rep(as.character(rng$name), 2), levels = stock_order))

d2$ssb_rel = d2$ssb / d2$msybtrigger
d3 = d2[is.finite(d2$ssb_rel), , drop = FALSE]
d3$name = factor(as.character(d3$name), levels = stock_order)

p1 = plot_advice_ssb(d1, pal, anchor = anchor)
p2 = plot_advice_ssb(d2, pal, anchor = anchor)
p3 = plot_advice_ssb(
  d3, pal,
  y            = "ssb_rel",
  ylab         = expression(SSB / B[trigger]),
  hline        = 1,
  status_bands = TRUE)

p1
p2
p3

save_beamer_fig(p1, "fig1.png", width = 8, height = 8.2)
save_beamer_fig(p2, "fig2.png", width = 8, height = 8.2)
save_beamer_fig(p3, "fig3.png", width = 8, height = 8.2)
```

```{r pel-1}
load(file.path(dirPlot, "ggPelForecas.RData"))

p4 = ggPelForecast + theme_beamer_ck() + xlab("Year")
p4
save_beamer_fig(p4, "fig4.png", width = 8, height = 4)
```

```{r dem-1}
load(file.path(dirPlot, "ggDemForecas.RData"))

p5 = ggDemForecast + theme_beamer_ck() + xlab("Year")
p5
save_beamer_fig(p5, "fig5.png", width = 8, height = 8)
```

```{r nep}
load(file.path(dirPlot, "ggNepForecastCatch.RData"))

p6 = ggNepForecastCatch + facet_wrap(~qname,ncol=2) + theme_beamer_ck() + xlab("Year")
p6
save_beamer_fig(p6, "fig6.png", width = 8, height = 8)
```

# Mackerel scenarios (fig_mac_*)

Status-quo F projections for `mac.27.nea`: catch under recruitment-level
shocks, a one-off M shock (×5 in 2035), and random recruitment noise.
All Beamer panels plot **catch** (not recruits). Helpers: `R/macProjections.R`.

```{r mac-params}
macSid      = "mac.27.nea"
macEndYr    = 2050L
macRecLevels = c(1, 0.75, 0.5, 0.25)
macMMul     = 5
macShockYr  = 2035L
macNiter    = 50L
macSeed     = 1L

# Catch panels need a visible y-axis (theme_beamer_ck hides axis.text.y).
theme_mac_catch = function(base_size = 12) {
  theme_beamer_ck(base_size = base_size) +
    theme(
      axis.text.y     = element_text(size = rel(0.85), colour = "grey20"),
      axis.title.y    = element_text(face = "bold", size = rel(1.05)),
      legend.position = "bottom")
}
```

```{r mac-om}
load(file.path(dirOM, "pel-om.RData"))

macStk  = stks[[macSid]]
macSr   = attributes(eqs[[macSid]])$sr
macFtar = fsqMean(macStk, n = 3L)

srsMac = as.data.frame(srs)
srsMac = srsMac[srsMac$sid == macSid, , drop = FALSE]
macRecBase = exp(mean(utils::tail(srsMac$rsd, 4)))
macRecSd   = stats::sd(srsMac$rsd, na.rm = TRUE)
```

```{r mac-rec, fig.width=8, fig.height=5}
macRecRuns = projectRecLevels(
  macStk, macSr, macFtar, macRecBase,
  levels = macRecLevels,
  endYr  = macEndYr)
dfMacRec = macProjToDf(macRecRuns)

ggMacRec = ggplot(dfMacRec, aes(year, Catch, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  labs(
    title = "Mackerel catch: different recruitment levels",
    subtitle = sprintf(
      "Fixed Fsq = %.3f; recruitment deviance = recentDev × level", macFtar),
    x = "Year", y = "Catch (t)") +
  theme_mac_catch()
ggMacRec
save_beamer_fig(ggMacRec, "fig_mac_rec.png", width = 8, height = 5)
```

```{r mac-mshock, fig.width=8, fig.height=5}
macMRuns = projectMShock(
  macStk, macSr, macFtar, macRecBase,
  mMul      = macMMul,
  shockYear = macShockYr,
  endYr     = macEndYr)
dfMacM = macProjToDf(macMRuns)

ggMacM = ggplot(dfMacM, aes(year, Catch, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  geom_vline(xintercept = macShockYr, linetype = 3, colour = "grey50") +
  labs(
    title = sprintf("Mackerel catch: one-off M × %s in %s", macMMul, macShockYr),
    subtitle = sprintf(
      "Fixed Fsq = %.3f; recruitment deviance = %.3f", macFtar, macRecBase),
    x = "Year", y = "Catch (t)") +
  theme_mac_catch()
ggMacM
save_beamer_fig(ggMacM, "fig_mac_mshock.png", width = 8, height = 5)
```

```{r mac-rand, fig.width=8, fig.height=5}
macRand = projectRandomRec(
  macStk, macSr, macFtar, macRecBase, macRecSd,
  endYr = macEndYr,
  niter = macNiter,
  seed  = macSeed)
dfMacRand = macProjToDf(macRand, scenarioCol = "iter")

macSumm = plyr::ddply(dfMacRand, "year", function(d) {
  data.frame(
    Catch_med = median(d$Catch),
    Catch_lo  = as.numeric(stats::quantile(d$Catch, 0.05)),
    Catch_hi  = as.numeric(stats::quantile(d$Catch, 0.95)))
})
macEx = dfMacRand[dfMacRand$iter == "iter1", , drop = FALSE]

ggMacRand = ggplot(macSumm, aes(year)) +
  geom_ribbon(
    aes(ymin = Catch_lo, ymax = Catch_hi),
    fill = "steelblue", alpha = 0.25) +
  geom_line(aes(y = Catch_med), colour = "steelblue", linewidth = 0.9) +
  geom_line(data = macEx, aes(y = Catch), colour = "black", linewidth = 0.5) +
  labs(
    title = "Mackerel catch: random recruitment noise",
    subtitle = sprintf(
      "Fixed Fsq = %.3f; %d iters; ribbon = 5–95%% catch; black = iter1",
      macFtar, macNiter),
    x = "Year", y = "Catch (t)") +
  theme_mac_catch()
ggMacRand
save_beamer_fig(ggMacRand, "fig_mac_rand.png", width = 8, height = 5)
```

## Backup metric panels (Catch, SSB, Recruits, F)

2×2 panels for each scenario family, written under `report/beamer/backup/`
(and `data/plot-objects/`). Beamer deck still uses the catch-only slides above.

```{r mac-backup-metrics, fig.width=10, fig.height=8}
dirMacBackup = file.path(dirBeamer, "backup")
dir.create(dirMacBackup, recursive = TRUE, showWarnings = FALSE)
dir.create(dirPlot, recursive = TRUE, showWarnings = FALSE)

save_mac_backup = function(plot, stem) {
  ggsave(file.path(dirMacBackup, paste0(stem, ".png")), plot,
         width = 10, height = 8)
  ggsave(file.path(dirPlot, paste0(stem, ".png")), plot,
         width = 10, height = 8)
  plot
}

ggMacRecMetrics = plotMacMetrics(
  dfMacRec,
  title = "Mackerel: recruitment levels",
  subtitle = sprintf("Fixed Fsq = %.3f; Catch / SSB / Recruits / F", macFtar))
ggMacRecMetrics
save_mac_backup(ggMacRecMetrics, "mac_rec_levels_metrics")

ggMacMMetrics = plotMacMetrics(
  dfMacM,
  title = sprintf("Mackerel: M × %s in %s", macMMul, macShockYr),
  subtitle = sprintf("Fixed Fsq = %.3f; Catch / SSB / Recruits / F", macFtar),
  vline = macShockYr)
ggMacMMetrics
save_mac_backup(ggMacMMetrics, "mac_mshock_metrics")

ggMacRandMetrics = plotMacMetricsStochastic(
  dfMacRand,
  title = "Mackerel: random recruitment noise",
  subtitle = sprintf(
    "Fixed Fsq = %.3f; %d iters; ribbon = 5–95%%; black = iter1",
    macFtar, macNiter))
ggMacRandMetrics
save_mac_backup(ggMacRandMetrics, "mac_random_rec_metrics")
```
