---
title: "Mackerel projections: recruitment, M shock, random rec"
subtitle: "`mac.27.nea` what-ifs on top of `pel-om.RData`"
author: "Laurence T. Kell"
date: "`r Sys.Date()`"
knit: (function(inputFile, encoding) { e <- new.env(); sys.source(file.path(dirname(normalizePath(inputFile)), "knit_to_report.R"), envir = e, keep.source = FALSE); e$knit_to_report(inputFile, encoding) })
output:
  html_document:
    self_contained: false
    lib_dir: "html/libs/06_mackerel"
    toc: true
    toc_float: true
    code_folding: show
---

# Purpose

Three illustrative projections for Northeast Atlantic mackerel
(`mac.27.nea`), all under **status-quo F** (mean Fbar of the last three
assessment years), starting from the bridged OM in `data/om/pel-om.RData`:

1. **Recruitment levels** — constant deviance multipliers (1, 0.75, 0.5, 0.25)
   relative to the recent residual mean.
2. **One-off M shock** — natural mortality × 5 for a single year
   (2035), then back to baseline M.
3. **Random recruitment** — year-to-year lognormal noise around the recent
   residual mean (example trajectories + median / 90% band).

Tune the parameters in the chunk below, then knit or source interactively.
Beamer panels (`fig_mac_*.png`) are exported from `report/05_figs.Rmd`.

---

```{r setup}
knitr::opts_chunk$set(warning = TRUE, message = TRUE, error = FALSE, fig.width = 9, fig.height = 5)
# fail-fast: do not swallow warnings (was options(warn = -1))
```

```{r paths}
source(file.path(dirname(knitr::current_input(dir = TRUE)), "ensureBimResilience.R"))
setupReportPaths()
loadReportLibraries(extra = "ggplot2")
```

```{r params}
sid      = "mac.27.nea"
endYr    = 2050L
fsqNyr   = 3L

# (i) recruitment multipliers applied to recent residual mean
recLevels = c(1, 0.75, 0.5, 0.25)

# (ii) one-off M pulse
mMul      = 5          # M × 5 for one year
shockYear = 2035L

# (iii) random recruitment
niter     = 50L
recSd     = NULL       # NULL → sd of historical log residuals
seed      = 1L
```

```{r load-om}
load(file.path(dirOM, "pel-om.RData"))

stk = stks[[sid]]
sr  = attributes(eqs[[sid]])$sr
ftar = fsqMean(stk, n = fsqNyr)

srsMac = as.data.frame(srs)
srsMac = srsMac[srsMac$sid == "mac.27.nea", , drop = FALSE]
recBase = exp(mean(utils::tail(srsMac$rsd, 4)))
if (is.null(recSd)) recSd = stats::sd(srsMac$rsd, na.rm = TRUE)

if (is.null(shockYear)) shockYear = dims(stk)$maxyear + 1L
btrig = rfpts$MSYBtrigger[rfpts$sid == "mac.27.nea"][1]

cat(sprintf(
  "mac.27.nea | maxyear=%s | Fsq=%.3f | recentDev=%.3f | recSd=%.3f | M shock year=%s (x%s)\n",
  dims(stk)$maxyear, ftar, recBase, recSd, shockYear, mMul))
```

# (i) Different recruitment levels

Constant recruitment deviance = `recentDev × level`, fixed status-quo F.

```{r rec-levels}
recRuns = projectRecLevels(
  stk, sr, ftar, recBase,
  levels = recLevels,
  endYr  = endYr)

dfRec = macProjToDf(recRuns)
dfRec$panel = "Recruitment levels"

ggRec = ggplot(dfRec, aes(year, Catch, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  labs(
    title = "Mackerel catch: different recruitment levels",
    subtitle = sprintf("Fixed Fsq = %.3f; deviance = recentDev × level", ftar),
    x = "Year", y = "Catch (t)") +
  theme_bw() +
  theme(legend.position = "bottom")
ggRec
```

# (ii) One-off M shock

Natural mortality × 5 in 2035, same F and baseline recruitment
deviance as the no-shock run. Plot shows **catch**.

```{r m-shock}
mRuns = projectMShock(
  stk, sr, ftar, recBase,
  mMul      = mMul,
  shockYear = shockYear,
  endYr     = endYr)

dfM = macProjToDf(mRuns)
dfM$panel = "M shock"

ggM = ggplot(dfM, aes(year, Catch, colour = Scenario)) +
  geom_line(linewidth = 0.7) +
  geom_vline(xintercept = shockYear, linetype = 3, colour = "grey50") +
  labs(
    title = sprintf("Mackerel catch: one-off M × %s in %s", mMul, shockYear),
    subtitle = sprintf("Fixed Fsq = %.3f; recruitment deviance = %.3f", ftar, recBase),
    x = "Year", y = "Catch (t)") +
  theme_bw() +
  theme(legend.position = "bottom")
ggM
```

# (iii) Random recruitment noise → catch

Year-to-year lognormal recruitment noise; plot shows resulting **catch**
(median, 90% band, one example path).

```{r random-rec}
stkRand = projectRandomRec(
  stk, sr, ftar, recBase, recSd,
  endYr = endYr,
  niter = niter,
  seed  = seed)

dfRand = macProjToDf(stkRand, scenarioCol = "iter")
dfRand$panel = "Random recruitment"

# Summary band + one example iteration
summ = plyr::ddply(dfRand, "year", function(d) {
  data.frame(
    Catch_med = median(d$Catch),
    Catch_lo  = as.numeric(stats::quantile(d$Catch, 0.05)),
    Catch_hi  = as.numeric(stats::quantile(d$Catch, 0.95)))
})

exIter = "iter1"
dfEx = subset(dfRand, iter == exIter)

ggRand = ggplot(summ, aes(year)) +
  geom_ribbon(aes(ymin = Catch_lo, ymax = Catch_hi), fill = "steelblue", alpha = 0.25) +
  geom_line(aes(y = Catch_med), colour = "steelblue", linewidth = 0.9) +
  geom_line(data = dfEx, aes(y = Catch), colour = "black", linewidth = 0.5) +
  labs(
    title = "Mackerel catch: random recruitment noise",
    subtitle = sprintf(
      "Fixed Fsq = %.3f; %d iters; ribbon = 5–95%% catch; black = %s",
      ftar, niter, exIter),
    x = "Year", y = "Catch (t)") +
  theme_bw()
ggRand
```

# Save outputs

```{r save}
dirMac = file.path(dirTAC, "mackerel")
dir.create(dirMac, recursive = TRUE, showWarnings = FALSE)
dir.create(dirPlot, recursive = TRUE, showWarnings = FALSE)

macProj = list(
  params = list(
    sid = sid, endYr = endYr, ftar = ftar, recBase = recBase, recSd = recSd,
    recLevels = recLevels, mMul = mMul, shockYear = shockYear,
    niter = niter, seed = seed),
  recLevels = recRuns,
  mShock    = mRuns,
  randomRec = stkRand,
  dfRec = dfRec,
  dfM   = dfM,
  dfRand = dfRand,
  summRand = summ)

save(macProj, file = file.path(dirOM, "mac-proj.RData"))

write.csv(dfRec,  file.path(dirMac, "mac-rec-levels.csv"), row.names = FALSE)
write.csv(dfM,    file.path(dirMac, "mac-m-shock.csv"),    row.names = FALSE)
write.csv(summ,   file.path(dirMac, "mac-random-rec-summary.csv"), row.names = FALSE)

ggsave(file.path(dirPlot, "ggMacRecLevels.png"),  ggRec,  width = 8, height = 5)
ggsave(file.path(dirPlot, "ggMacMShock.png"),     ggM,    width = 8, height = 5)
ggsave(file.path(dirPlot, "ggMacRandomRec.png"),  ggRand, width = 8, height = 5)

# Backup 2×2 panels: Catch, SSB, Recruits, F
dirMacBackup = file.path(dirReport, "beamer", "backup")
dir.create(dirMacBackup, recursive = TRUE, showWarnings = FALSE)

ggRecMetrics = plotMacMetrics(
  dfRec,
  title = "Mackerel: recruitment levels",
  subtitle = sprintf("Fixed Fsq = %.3f; Catch / SSB / Recruits / F", ftar))
ggMMetrics = plotMacMetrics(
  dfM,
  title = sprintf("Mackerel: M × %s in %s", mMul, shockYear),
  subtitle = sprintf("Fixed Fsq = %.3f; Catch / SSB / Recruits / F", ftar),
  vline = shockYear)
ggRandMetrics = plotMacMetricsStochastic(
  dfRand,
  title = "Mackerel: random recruitment noise",
  subtitle = sprintf(
    "Fixed Fsq = %.3f; %d iters; ribbon = 5–95%%; black = %s",
    ftar, niter, exIter))

ggsave(file.path(dirMacBackup, "mac_rec_levels_metrics.png"), ggRecMetrics,
       width = 10, height = 8)
ggsave(file.path(dirMacBackup, "mac_mshock_metrics.png"), ggMMetrics,
       width = 10, height = 8)
ggsave(file.path(dirMacBackup, "mac_random_rec_metrics.png"), ggRandMetrics,
       width = 10, height = 8)
ggsave(file.path(dirPlot, "mac_rec_levels_metrics.png"), ggRecMetrics,
       width = 10, height = 8)
ggsave(file.path(dirPlot, "mac_mshock_metrics.png"), ggMMetrics,
       width = 10, height = 8)
ggsave(file.path(dirPlot, "mac_random_rec_metrics.png"), ggRandMetrics,
       width = 10, height = 8)
```

## Summary

| Scenario | What varies | Held fixed | Beamer | Backup |
|----------|-------------|------------|--------|--------|
| Recruitment levels | Constant deviance multiplier | Fsq, M | Catch | Catch/SSB/Rec/F |
| M shock | M × `r mMul` in `r shockYear` only | Fsq, baseline deviance | Catch | Catch/SSB/Rec/F |
| Random recruitment | Year-to-year lognormal noise | Fsq, mean deviance | Catch | Catch/SSB/Rec/F |

Artefacts: `data/om/mac-proj.RData`, `data/TAC/mackerel/*.csv`,
`data/plot-objects/`, `report/beamer/backup/mac_*_metrics.png`.

Beamer catch panels (`fig_mac_*.png`) are written by `report/05_figs.Rmd`.
